---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## 1. Loading and preprocessing the data
```{r echo=TRUE}
data<-read.csv("C:/Users/MAHE/Desktop/Career/RepData_PeerAssessment1/activity/activity.csv")
data$date <- as.Date(data$date, format = "%Y-%m-%d")
```


## 2. What is mean total number of steps taken per day?
### Total Number of Daily Steps
```{r}
library(ggplot2)
daily_steps<-tapply(data[,1],data[,2],sum,na.rm=T)
```
### Histogram of the total number of steps taken each day
```{r}
qplot(daily_steps,binwidth=2000,xlab="Total Steps Taken Per Day")
```

### Mean and Median of the total number of steps taken per day 
```{r}
mean(daily_steps) 
median(daily_steps)

```


## 3. What is the average daily activity pattern?
### Time series plot of the 5-minute interval and the average number of steps
```{r}
interval_steps<-tapply(data[,1],data[,3],mean,na.rm=T)
plot(interval_steps,type="l",xlab="5-minute Interval",ylab="Average Steps")
```

### 5-miute interval which on average contain maximum number of steps

```{r}
as.numeric(names(which.max(interval_steps)))
```


## 4. Imputing missing values
###  The total number of missing values in the dataset
```{r}
 sum(!complete.cases(data))
```
### Device Strategy to Impute Missing Values
 We will replace the Missing Values with respective 5-minute Interval Steps Value

### Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
library(dplyr)
imputed_data<-data %>% group_by(interval) %>% mutate(steps=ifelse(is.na(steps),mean(steps,na.rm=T),steps))
imputed_data<-data.frame(imputed_data)
```
### Histogram, Mean and Median of Imputed Datasets

```{r}
imputed_daily_steps<-tapply(imputed_data[,1],imputed_data[,2],sum)
qplot(imputed_daily_steps,binwidth=2000,xlab="Total Steps Taken Per Day")
mean(imputed_daily_steps)
median(imputed_daily_steps)

```

## 5. Are there differences in activity patterns between weekdays and weekends?

### Creating a factor Variable with two level 'Weekdays' and 'Weekends'

```{r}
 days<-ifelse(weekdays(imputed_data$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
 days<-as.factor(days)
 imputed_data<-cbind(imputed_data,days)
    ```
 
###  A panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend day
 
```{r}
x<-imputed_data %>% group_by(days,interval) %>% summarise(steps=sum(steps))
x<-data.frame(x)
qplot(interval,steps,data=x,geom="line",facets = days~.)
```
